\section{Materials and Methods}

\subsection{Dataset and Feature Engineering}

The dataset simulates public transport in a controlled environment~\cite{ribeiro2026eucnc}. Zone~A represents the vehicle interior with an access point near the door. Zone~B represents the bus stop in an adjacent corridor. The same four devices used in the latest experiment \cite{ribeiro2026eucnc} execute four movement patterns, denoted AA (inside vehicle), BB (at stop), BA (boarding), and AB (alighting). Each device produces 10 RSSI readings at 1~s intervals per observation.

Two scenarios were collected. The pure scenario uses a single device with no interference (160 samples). The noise scenario uses four concurrent devices executing different route combinations (1,200 samples). The combined dataset contains 2,253 labeled samples distributed as AA 578, AB 550, BA 546, and BB 577. The noise-only subset contains 1,196 samples (AA 298, AB 300, BA 298, BB 300).

Four engineered channels are extracted from the raw RSSI sequence $r_t$. The raw RSSI $r_t$ provides absolute signal position. Velocity $\Delta_t = r_t - r_{t-1}$ captures movement speed. Acceleration $\Delta^2_t$ captures rate of change. Window deviation $d_t = r_t - \mu_w$, where $\mu_w$ is the per-sample mean, captures trajectory shape independent of absolute RSSI level. These channels yield input tensors of shape $(N, 10, 4)$.

\subsection{Proposed Architectures}

Table~\ref{tab:architectures} lists the 11 evaluated architectures. All models output 4-class logits through a linear classification head.

\begin{table}[H]
    \centering
    \caption{DL Architectures Evaluated}
    \label{tab:architectures}
    \resizebox{\columnwidth}{!}{\begin{tabular}{lll}\toprule
            Architecture                        & Type               & Key characteristic                                         \\\midrule
            RNN, GRU, LSTM, BiLSTM              & Recurrent          & 2 layers, hidden 64                                        \\
            TCN~\cite{bai2018tcn}               & Convolutional      & Dilated causal, residual                                   \\
            Transformer                         & Attention          & 4-head self-attention, sinusoidal PE                       \\
            MambaClassifier~\cite{gu2026mamba3} & State-space        & $O(L)$ complexity, rank-$R$=16                             \\
            CNN2DRNN                            & Hybrid 2D+RNN      & Conv2d $(3,3)$,$(5,3)$ $\to$ LSTM/GRU + attention          \\
            MetaFusion                          & Context-aware      & LSTM/GRU fused with learned vector representations (dim~8) \\
            Multi-Scale CNN                     & Multi-branch       & Kernels $(3,5,7)$ + residual + attention                   \\
            SoftMoE                             & Mixture of Experts & Heterogeneous experts, learned top-2 gating routing        \\
            Autoformer~\cite{wu2021autoformer}  & Decomposition      & FFT auto-correlation, trend-seasonal blocks                \\
            DeepStackEnsemble                   & Two-stage meta     & 18 bases $\to$ per-class Level2Net $\to$ residual MLP      \\\bottomrule
        \end{tabular}}
\end{table}

The DeepStackEnsemble uses per-class Level2Nets: small MLPs receiving concatenated logits from the base models and outputting 4-class logits. Table~\ref{tab:ds_bases} lists the 18 base models (2 configurations per architecture) that feed the ensemble.

\begin{table}
    \centering
    \caption{DeepStack Base Models}
    \label{tab:ds_bases}
    \begin{tabular}{lll}\toprule
        Architecture & Config A (light) & Config B (heavy) \\\midrule
        RNN          & dim=64, L=2, d=0.2       & dim=256, L=4, d=0.3, bidir  \\
        GRU          & dim=64, L=2, d=0.2       & dim=256, L=3, d=0.3, bidir  \\
        LSTM         & dim=64, L=2, d=0.2       & dim=256, L=3, d=0.3, bidir  \\
        CNN          & dim=64, blk=2, k=[3,5]     & dim=256, blk=4, k=[3,5,7,9]  \\
        TCN          & dim=128, d=3, k=4          & dim=512, d=7, k=8            \\
        Transformer  & dim=64, H=4, L=2           & dim=256, H=8, L=4            \\
        BiLSTM       & dim=64, L=2, d=0.2        & dim=256, L=3, d=0.3          \\
        CNN2DRNN     & Conv2d $\to$ GRU, d=0.2   & Conv2d $\to$ LSTM, d=0.3     \\
        Mamba        & d\_model=64, L=2           & d\_model=256, L=4            \\\bottomrule
    \end{tabular}
\end{table}

\subsection{Augmentation and Implementation}

Two strategies expand the training data from 2,253 to 10,115 samples. Class-conditioned Gaussian noise adds 6 copies to clean samples ($\sigma=3.0$~dBm) and 4 copies to noisy samples ($\sigma=0.8$~dBm). Intra-class Mixup blends 600 random pairs per class as $X_{\text{new}} = \lambda X_1 + (1-\lambda) X_2$ with $\lambda \sim \text{Beta}(0.4,0.4)$, restricted to same-label pairs.

The pipeline uses PyTorch~2.6 and Optuna~4.7. Training uses AdamW optimizer with cosine annealing, early stopping (patience 25), label smoothing ($\varepsilon=0.1$), Mixup augmentation ($\alpha=0.2$), L1 regularization ($\lambda=10^{-5}$), L2 weight decay ($10^{-4}$), dropout ($p=0.3$), gradient clipping at 1.0, and batch size 64. HPO uses 50 Optuna TPE trials per model. SoftMoE additionally uses NSGA-II multi-objective optimization. The primary metric is MCC~\cite{chicco2020mcc}. All experiments use 5-fold stratified cross-validation with 80/20 splits and 3 random seeds (3, 5, 42). A total of 65 models were evaluated.
