% Related Work — tightened, better flow
\section{Related Work}

Passenger counting spans multiple sensing modalities. Video-based APC using YOLOv5 achieves high accuracy but degrades under occlusion and variable lighting~\cite{pronello2023evaluating}; edge AI cameras address scalability at the cost of deployment complexity~\cite{ono2025realtime}. ML pipelines on traditional APC data enable offline correction of sensor errors~\cite{barabino2014offline,wiboonsiriruk2023efficient}. All camera-based methods raise privacy concerns and require substantial infrastructure.

Wi-Fi sensing avoids these drawbacks. iABACUS~\cite{nitti2020iabacus} and Myrvoll et al.~\cite{myrvoll2017counting} demonstrated counting via probe requests and signature analysis. Fabre et al.~\cite{fabre2025machine} applied RF and gradient boosting for ridership estimation. Oransirikul et al.~\cite{oransirikul2019classifying} distinguished passenger from non-passenger activity, while Simoncic et al.~\cite{simoncic2023nonintrusive} monitored presence through passive probes. CSI-based methods~\cite{wang2024csi,guo2024rssi} offer finer signal characterization but depend on specialized hardware unavailable in commodity APs. RSSI fingerprinting for localization has been extensively surveyed~\cite{singh2021rssisurvey,zholamanov2025rssi,agualimpia2024rssi}; BLE beacons~\cite{jain2021lowcost} are an alternative but require dedicated infrastructure. Cerqueira et al.~\cite{cerqueira2022inference} addressed the complementary problem of OD matrix inference from smart card data.

Recent DL advances inform our architectural choices. Mamba~\cite{gu2023mamba} introduced selective state-space models with $O(L)$ complexity, extended by Mamba-2~\cite{dao2024mamba2} and Mamba-3 MIMO~\cite{gu2026mamba3}. Autoformer~\cite{wu2021autoformer} replaced self-attention with auto-correlation for time series. Mixture of Experts~\cite{shazeer2017moe} enables conditional computation through sparse activation. Our prior work~\cite{ribeiro2026eucnc} established classical ML baselines (MCC~$=$~0.756) on this dataset across 34 algorithms. This paper extends that analysis with 11$+$ specialized DL architectures, comprehensive HPO via Optuna~\cite{optuna2019}, and systematic interference analysis.
